---
title: "Marketing Campaign EDA"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
```

```{r}
mkt <- read.csv("marketing_campaign.csv", sep = ';')
```

```{r}
str(mkt)
```

## Data Cleaning -- Missing Values
```{r}
income_na <- sum(is.na(mkt$Income))
income_na
```

```{r}
mkt$Income <- ifelse(is.na(mkt$Income), round(mean(mkt$Income, na.rm = TRUE), 0), mkt$Income)
```

```{r}
income_no_na <- sum(is.na(mkt$Income))
income_no_na
```

## Data Preparation -- Variable Manipulation
```{r}
mkt$previous_response <- mkt$AcceptedCmp1+mkt$AcceptedCmp2+mkt$AcceptedCmp3+mkt$AcceptedCmp4+ mkt$AcceptedCmp5
```

```{r}
library(lubridate)
```

```{r}
mkt$year = substr(mkt$Dt_Customer, 1, 4)
mkt$date_now <- today()
mkt$year_now = substr(mkt$date_now, 1, 4)
mkt$tenure = as.numeric(mkt$year_now) - as.numeric(mkt$year)
mkt$Age = as.numeric(mkt$year_now) - as.numeric(mkt$Year_Birth)
```

```{r}
mkt_new <- mkt[ , c(3:7,9:30, 34:35)]
```

```{r}
unique(mkt_new$Marital_Status)
```

```{r}
Single <- c('Single')
Married <- c("Together", "Married")
Alone <- c("Divorced", "Widow", "Alone")
Not_sure <- c("Absurd", "YOLO")
```

```{r}
mkt_new$Marital_Status = sapply(as.vector(mkt_new$Marital_Status), function(x) if(x %in% Single) "Single" else x )
mkt_new$Marital_Status = sapply(as.vector(mkt_new$Marital_Status), function(x) if(x %in% Married) "Married" else x )
mkt_new$Marital_Status = sapply(as.vector(mkt_new$Marital_Status), function(x) if(x %in% Alone) "Alone" else x )
mkt_new$Marital_Status = sapply(as.vector(mkt_new$Marital_Status), function(x) if(x %in% Not_sure) "Not_sure" else x )
```

```{r}
unique(mkt_new$Marital_Status)
```

## 1D Marginal Distribution
```{r}
library(ggplot2)
library(gridExtra)
```

## Bar Chart
```{r}
plot_education <- ggplot(data = mkt_new, aes(x = factor(Education), fill = factor(Education))) +
  geom_bar() +
  ggtitle("Education") +
  guides(fill=FALSE)+
  xlab("Education") 

plot_Marital_Status <- ggplot(data = mkt_new, aes(x = factor(Marital_Status), fill = factor(Marital_Status))) +
  geom_bar() +
  ggtitle("Marital Status") +
  guides(fill=FALSE)+
  xlab("Marital Status") 

plot_Kidhome <- ggplot(data = mkt_new, aes(x = factor(Kidhome), fill = factor(Kidhome))) +
  geom_bar() +
  ggtitle("Kidhome") +
  guides(fill=FALSE)+
  xlab("Kidhome") 

plot_Teenhome <- ggplot(data = mkt_new, aes(x = factor(Teenhome), fill = factor(Teenhome))) +
  geom_bar() +
  ggtitle("Teenhome") +
  guides(fill=FALSE)+
  xlab("Teenhome") 

grid.arrange(plot_education, plot_Marital_Status, plot_Kidhome, plot_Teenhome, ncol=2)
```

## NumDealsPurchases
```{r}
unique(mkt_new$NumDealsPurchases)
```

## NumWebPurchases
```{r}
unique(mkt_new$NumWebPurchases)
```

## NumCatalogPurchases
```{r}
unique(mkt_new$NumCatalogPurchases)
```

## NumStorePurchases
```{r}
unique(mkt_new$NumStorePurchases)
```

```{r}
plot_NumDealsPurchases <- ggplot(data = mkt_new, aes(x = factor(NumDealsPurchases), fill = factor(NumDealsPurchases))) +
  geom_bar() +
  ggtitle("Number of DealsPurchases") +
  guides(fill=FALSE)+
  xlab("Number of DealsPurchases") 

plot_NumWebPurchases <- ggplot(data = mkt_new, aes(x = factor(NumWebPurchases), fill = factor(NumWebPurchases))) +
  geom_bar() +
  ggtitle("Number of WebPurchases") +
  guides(fill=FALSE)+
  xlab("Number of WebPurchases") 

plot_NumCatalogPurchases <- ggplot(data = mkt_new, aes(x = factor(NumCatalogPurchases), fill = factor(NumCatalogPurchases))) +
  geom_bar() +
  ggtitle("Number of CatalogPurchases") +
  guides(fill=FALSE)+
  xlab("Number of CatalogPurchases") 

plot_NumStorePurchases <- ggplot(data = mkt_new, aes(x = factor(NumStorePurchases), fill = factor(NumStorePurchases))) +
  geom_bar() +
  ggtitle("Number of StorePurchases") +
  guides(fill=FALSE)+
  xlab("Number of StorePurchases") 

grid.arrange(plot_NumDealsPurchases, plot_NumWebPurchases, plot_NumCatalogPurchases, plot_NumStorePurchases, ncol=2)
```

## Marital Status
```{r}
ms = mkt_new$Marital_Status
as.data.frame(sort(table(ms),decreasing = TRUE))
```

## Histogram
```{r}
plot_Income <- ggplot(mkt_new, aes(x=log(Income), fill="#FF6666")) +
  geom_histogram(color="white",alpha=0.7) +
  labs(title="Income") +
  xlab("Income") +
  theme(legend.position="none", plot.title=element_text(hjust=0.5))

plot_Recency <- ggplot(mkt_new, aes(x=Recency)) +
  geom_histogram(color="white",alpha=0.7, fill="darkgreen") +
  labs(title="Recency") +
  xlab("Recency") +
  theme(legend.position="none", plot.title=element_text(hjust=0.5))

plot_MntWines <- ggplot(mkt_new, aes(x=log(mkt_new$MntWines+1))) +
  geom_histogram(color="white",alpha=0.7, fill="orange") +
  labs(title="Amount of Wines Purchased") +
  xlab("Amount of Wines Purchased") +
  theme(legend.position="none", plot.title=element_text(hjust=0.5))

plot_MntFruits <- ggplot(mkt_new, aes(x=log(mkt_new$MntFruits))) +
  geom_histogram(color="white",alpha=0.7, fill="blue") +
  labs(title="Amount of Fruits Purchased") +
  xlab("Amount of Fruits Purchased") +
  theme(legend.position="none", plot.title=element_text(hjust=0.5))

grid.arrange(plot_Income, plot_Recency, plot_MntWines, plot_MntFruits, ncol=2)
```

```{r}
plot_Meat <- ggplot(mkt_new, aes(x=log(mkt_new$MntMeatProducts))) +
  geom_histogram(color="white",alpha=0.7, fill="dark red") +
  labs(title="Amount of Meat Purchased") +
  xlab("Amount of Meat Purchased") +
  theme(legend.position="none", plot.title=element_text(hjust=0.5))

plot_fish <- ggplot(mkt_new, aes(x=log(mkt_new$MntFishProducts))) +
  geom_histogram(color="white",alpha=0.7, fill="dark orange") +
  labs(title="Amount of Fish Purchased") +
  xlab("Amount of Fish Purchased") +
  theme(legend.position="none", plot.title=element_text(hjust=0.5))

plot_Sweet <- ggplot(mkt_new, aes(x=log(mkt_new$MntSweetProducts))) +
  geom_histogram(color="white",alpha=0.7, fill="purple") +
  labs(title="Amount of Sweet Products Purchased") +
  xlab("Amount of Sweet Products Purchased") +
  theme(legend.position="none", plot.title=element_text(hjust=0.5))

plot_Gold <-ggplot(mkt_new, aes(x=log(mkt_new$MntGoldProds+1))) +
  geom_histogram(color="white",alpha=0.7, fill="pink") +
  labs(title="Amount of Gold Purchased") +
  xlab("Amount of Gold Purchased") +
  theme(legend.position="none", plot.title=element_text(hjust=0.5))

grid.arrange(plot_Meat, plot_fish, plot_Sweet, plot_Gold, ncol=2)
```

## Previous Campaign Response Rate
```{r}
cmp2 = mkt_new$AcceptedCmp2
as.data.frame(sort(table(cmp2),decreasing = TRUE))
```

```{r}
p1 <- ggplot(data = mkt_new, aes(x = factor(AcceptedCmp1), fill = factor(AcceptedCmp1))) +
  geom_bar() +
  ggtitle("AcceptedCmp1") +
  guides(fill=FALSE)+
  xlab("AcceptedCmp1") 
p11 <- p1 + scale_fill_brewer(palette="Paired") + theme_minimal()

p2 <- ggplot(data = mkt_new, aes(x = factor(AcceptedCmp2), fill = factor(AcceptedCmp2))) +
  geom_bar() +
  ggtitle("AcceptedCmp2") +
  guides(fill=FALSE)+
  xlab("AcceptedCmp2") 
p21 <- p2 + scale_fill_brewer(palette="Paired") + theme_minimal()

p3 <- ggplot(data = mkt_new, aes(x = factor(AcceptedCmp3), fill = factor(AcceptedCmp3))) +
  geom_bar() +
  ggtitle("AcceptedCmp3") +
  guides(fill=FALSE)+
  xlab("AcceptedCmp3") 
p31 <- p3 + scale_fill_brewer(palette="Paired") + theme_minimal()

p4 <- ggplot(data = mkt_new, aes(x = factor(AcceptedCmp4), fill = factor(AcceptedCmp4))) +
  geom_bar() +
  ggtitle("AcceptedCmp4") +
  guides(fill=FALSE)+
  xlab("AcceptedCmp4") 
p41 <- p4 + scale_fill_brewer(palette="Paired") + theme_minimal()

p5 <- ggplot(data = mkt_new, aes(x = factor(AcceptedCmp5), fill = factor(AcceptedCmp5))) +
  geom_bar() +
  ggtitle("AcceptedCmp5") +
  guides(fill=FALSE)+
  xlab("AcceptedCmp5") 
p51 <- p5 + scale_fill_brewer(palette="Paired") + theme_minimal()

grid.arrange(p11, p21, p31, p41, p51, ncol=2)
```

## Previous Response in Total
```{r}
ggplot(data = mkt_new, aes(x = factor(previous_response), fill = factor(previous_response))) +
  geom_bar() +
  ggtitle("previous_response") +
  guides(fill=FALSE)+
  xlab("previous_response") 
```

## Current Response
```{r}
ggplot(data = mkt_new, aes(x = factor(Response), fill = factor(Response))) +
  geom_bar() +
  ggtitle("Response") +
  guides(fill=FALSE)+
  xlab("Response") 
```

## Age
```{r}
ggplot(mkt_new, aes(x=log(mkt_new$Age))) +
  geom_histogram(color="white",alpha=0.7, fill="dark green") +
  labs(title="Age") +
  theme(legend.position="none", plot.title=element_text(hjust=0.5))
```

## tenure
```{r}
ggplot(data = mkt_new, aes(x = factor(tenure), fill = factor(tenure))) +
  geom_bar() +
  ggtitle("tenure") +
  guides(fill=FALSE)+
  xlab("tenure") 
```

## 2D Correlation Plot
```{r}
library(vcd)
```

## Education
```{r}
mosaic(Response ~ Education, data = mkt_new)
```

## Marital Status
```{r}
mosaic(Response ~ Marital_Status, data = mkt_new)
```

## Previous Response
```{r}
mosaic(Response ~ previous_response, data = mkt_new)
```

## tenure
```{r}
mosaic(Response ~ tenure, data = mkt_new)
```

## Box Plot
```{r}
bp2 <- ggplot(data = mkt_new, aes(y=Recency, x = factor(Response), fill = factor(Response))) +
  geom_boxplot() +
  xlab("Response") +
  ylab("Recency") +
  ggtitle("Response VS Recency")
c1 <- bp2 + scale_fill_brewer(palette="Paired") + theme_minimal()

c2 <- ggplot(data = mkt_new, aes(y=log(MntWines), x = factor(Response), fill = factor(Response))) +
  geom_boxplot() +
  xlab("Response") +
  ylab("MntWines") +
  ggtitle("Response VS MntWines")

c3 <-ggplot(data = mkt_new, aes(y=log(mkt_new$MntFruits), x = factor(Response), fill = factor(Response))) +
  geom_boxplot() +
  xlab("Response") +
  ylab("MntFruits") +
  ggtitle("Response VS MntFruits")

c4 <-ggplot(data = mkt_new, aes(y=log(mkt_new$MntMeatProducts), x = factor(Response), fill = factor(Response))) +
  geom_boxplot() +
  xlab("Response") +
  ylab("MntMeatProducts") +
  ggtitle("Response VS MntMeatProducts")

c5 <-ggplot(data = mkt_new, aes(y=log(mkt_new$MntFishProducts), x = factor(Response), fill = factor(Response))) +
  geom_boxplot() +
  xlab("Response") +
  ylab("MntFishProducts") +
  ggtitle("Response VS MntFishProducts")

c6 <-ggplot(data = mkt_new, aes(y=log(mkt_new$MntGoldProds), x = factor(Response), fill = factor(Response))) +
  geom_boxplot() +
  xlab("Response") +
  ylab("MntGoldProds") +
  ggtitle("Response VS MntGoldProds")

grid.arrange(c1, c2, c3, c4, c5, c6, ncol=2)
```

## Age
```{r}
bp3 <- ggplot(data = mkt_new, aes(y=log(mkt_new$Age), x = factor(Response), fill = factor(Response))) +
  geom_boxplot() +
  xlab("Response") +
  ylab("Age") +
  ggtitle("Response VS Age")
bp3 + scale_fill_brewer(palette="BuGn") + theme_minimal()
```


